Single-cell virus sequencing of influenza variants that trigger innate immunity

### 2018-09-10 paper/paper.tex
revised methods for viruses for single-cell

\subsection{Viruses for single-cell experiments}
\jdbcomment{Alistair writes this}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.

@@

\subsection{Viruses for validation experiments}
\jdbcomment{Alistair writes this}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.

>>>>

\subsection{Viruses for single-cell experiments}
We performed the single-cell experiments using the A/WSN/1933 (H1N1) strain of influenza virus.
The cells were infected with a roughly equal mix of the wildtype virus and a variant of the virus where synonymous mutations were added within a few 100 nucleotides of each termini of each gene segment.
We have used a similar synonymous viral barcoding strategy in our prior single-cell work~\citep{russell2018extreme} as it allows us to detect about half of co-infected cells based on the expression of both viral barcode variants.
In the current work, we extended this approach by placing synonymous barcodes near \emph{both} termini of the gene segments in order to allow us to quantify strand exchange during PacBio sequencing (\FIGSUPP[genotypes]{StrandExchange}).
The sequences of all gene segments from the wildtype and synonymously barcoded viral strains are in \FIGDATA[experiment]{virus_seqs}.

Both viral strains were generated by reverse genetics using the pHW18* series of bi-directional plasmids~\citep{hoffmann2000dna}.
We controlled the durations and MOI during viral passaging since these factors can greatly affect the accumulation of defective viral particles~\citep{xue2016propagation}.
The viruses were generated by reverse genetics in co-cultures of 293T and MDCK-SIAT1 in influenza growth media and then propagated in MDCK-SIAT1 cells in influenza growth media using the same basic procedures detailed in \citet{russell2018extreme}.
Specifically, after generation by reverse genetics, the wildtype variant was expanded at an MOI of 0.001 for 72 hours twice in MDCK-SIAT1 cells, and the synonymously barcoded variant was expanded once at an MOI of 0.01 for 60 hours.
The MOIs for this passaging are based on titers determined using TCID50 assays via the formula of \citet{reed1938simple} as implemented at \url{https://github.com/jbloomlab/reedmuenchcalculator}.

For the single-cell experiments (which only examine the transcriptional results of a single cycle of infection), we were most interested in the titer of viral particles that are transcriptionally active for a single round of infection of A549 cells.
We estimated titers of transcriptionally active virions by staining for HA expression in virus-infected A549 cells.
Specifically, we infected A549 cells in influenza growth medium, and at \jdbcomment{?} hours post-infection, we trypsinized cells, re-suspended in phosphate-buffered saline (PBS) supplemented with 2\% heat-inactivated fetal bovine serum (FBS), and stained with 10 $\mu$g/ml of H17-L19, a mouse monoclonal antibody previously shown to bind to the HA from the A/WSN/1933 strain of virus~\citep{doud2017complete}.
After washing in PBS supplemented with 2\% FBS, the cells were stained with a goat anti-mouse IgG antibody conjugated to APC, washed, fixed in 1\% formaldehyde in PBS, washed again, and then analyzed by flow cytometry to determine the fraction expressing detectable HA protein.

@@

\subsection{Viruses for validation experiments}
\jdbcomment{Alistair writes this. I've copied text from prior consolidated section, but it needs to be clarified by more clearly breaking things down by virus.}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.
